RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-09-16 doc/paper.tex
Intro: drafted

	 \begin{table}[!ht]
		\centering
		\scalebox{0.7}{
		\begin{tabularx}{450pt}{l  l  X}  
			Algorithm          &   Paper/year &   Short details   \\ \hline
	        Hybrid GP  &   \cite{raidl1998hybrid} (1998)   & GP--based method; solutions locally optimizes by method of least squares to find optimum coefficients for top-level terms \\ 
	        DFE        & \cite{cerny2008using} (2008)   & Differential evolution algorithm \\
	        Eurequa                & \cite{schmidt2009distilling, schmidt2011machine} (2009, 2011) & Age-fitness pareto optimization approach using co-evolved fitness estimation   \\
        	APF            &      \cite{schmidt2010age} (2010)        &  Age-fitness pareto optimization approach                        \\
        	FFX  & \cite{mcconaghy2011ffx} (2011)  & The Fast Function Extraction algorithm --  non-evolutionary technique based on a machine learning technique called path-wise regularized learning \\
             ABCP &  \cite{karaboga2012artificial} (2012)  & Artificial bee colony programming  approach \\
			EPLEX          &      \cite{la2016epsilon} (2016)         &   A parent selection method called $\epsilon$--lexicase selection  \\ 
	       Local optimization     & \cite{kommenda2018local} (2018)   & Constants Optimization in GP by Nonlinear Least Square \\
	       SBP-GP &  \cite{virgolin2019linear} (2019)  & the idea of semantic back-propagation \\
	       BSR & \cite{jin2019bayesian} (2019) & ML--based approach; Bayesian symbolic regression  \\ 
	       DSR  & \cite{petersen2019deep} (2019) & Deep Symbolic Regression based on a RNN approach further utilizing the policy gradient search \\
	       OPERON & \cite{kommenda2020parameter} (2020) &  Utilizing nonlinear least squares for parameter identification of SR models with LS \\
	       VNP    & \cite{elleuch2020variable} (2020) & VNS based GP approach \\
	       OccamNet & \cite{costa2020fast} (2020) &   A fast neural network approach; the model defines a probability distribution over a non-differentiable function space; it samples functions and updates the weights with back-propagation  based on cross-entropy matching in an EA strategy	 \\

@@

	The progress in the SR field suffered from a lack of uniform, robust, and transparent
	benchmarking standards over the last two decade. Recently, La Cava et al.~\cite{la2021contemporary} proposed an
	open-source, reproducible benchmarking platform for SR, called SRBench. The authors extended  PMLB~\cite{olson2017pmlb},  a repository of standardized regression by 130 SR datasets for which model forms are known, see more in the aforementioned paper.\fxnote{TODO: Describe results of experimental evaluation}
\subsection{Contributions of this work}

The main contributions of this work can be summarized as follows:

@@

	\item The new set of unbiased instances is introduced -- randomly generated formula of various sizes. All the other methods were tested against this new test-bed. The proposed RILS-ROLS method showed to be dominant here as well. 
\end{enumerate}

>>>>

	 \begin{table}[!ht]
		\centering
		\scalebox{0.7}{
		\begin{tabularx}{450pt}{l  l  X}  
			Algorithm          &   Paper/year &   Short details   \\ \hline
	        Hybrid GP  &   \cite{raidl1998hybrid} (1998)   & GP--based method; solutions locally optimizes by method of least squares to find optimum coefficients for top-level terms \\ 
	        DFE        & \cite{cerny2008using} (2008)   & Differential evolution algorithm \\
	        Eurequa                & \cite{schmidt2009distilling, schmidt2011machine} (2009, 2011) & Age-fitness pareto optimization approach using co-evolved fitness estimation   \\
        	APF            &      \cite{schmidt2010age} (2010)        &  Age-fitness pareto optimization approach                        \\
        	FFX  & \cite{mcconaghy2011ffx} (2011)  & The Fast Function Extraction algorithm --  non-evolutionary technique based on a machine learning technique called path-wise regularized learning \\
             ABCP &  \cite{karaboga2012artificial} (2012)  & Artificial bee colony programming  approach \\
			EPLEX          &      \cite{la2016epsilon} (2016)         &   A parent selection method called $\epsilon$--lexicase selection  \\ 
	       Local optimization     & \cite{kommenda2018local} (2018)   & Constants Optimization in GP by Nonlinear Least Square \\
	       SBP-GP &  \cite{virgolin2019linear} (2019)  & The idea of semantic back-propagation utilized in GP \\
	       BSR & \cite{jin2019bayesian} (2019) & ML--based approach; Bayesian symbolic regression  \\ 
	       DSR  & \cite{petersen2019deep} (2019) & Deep Symbolic Regression based on a RNN approach further utilizing the policy gradient search \\
	       OPERON & \cite{kommenda2020parameter} (2020) &  Utilizing nonlinear least squares for parameter identification of SR models with LS \\
	       VNP    & \cite{elleuch2020variable} (2020) & VNS based GP approach \\
	       OccamNet & \cite{costa2020fast} (2020) &   A fast neural network approach; the model defines a probability distribution over a non-differentiable function space; it samples functions and updates the weights with back-propagation  based on cross-entropy matching in an EA strategy	 \\

@@

	The progress in the SR field suffered from a lack of uniform, robust, and transparent
	benchmarking standards over the last two decade. Recently, La Cava et al.~\cite{la2021contemporary} proposed an
	open-source, reproducible benchmarking platform for SR, called SRBench. The authors extended  PMLB~\cite{olson2017pmlb},  a repository of standardized regression, by 130 SR datasets for which model forms are known; see more in the aforementioned paper. In the extensive experimental evaluation  where 14
	symbolic regression methods and 7 machine learning methods are compared on the  set of 252 diverse
    regression problems. The general conclusions derived from there are: ($i$) when comes to real-world problem, the best performing algorithms are those which combine genetic algorithm with parameter estimation; ($ii$) when come to the instances with the presence of noise, the GP--based and deep learning techniques are performing similarly.

    In this work we present yet another GP--based approach to solve SR, which utilizes the popular iterated local search meta-heuristic~\cite{lourencco2003iterated,lourencco2019iterated}; in the remaining text, we label the metod by \textsc{Rils-Rols}.  The main code of the efficient local search is utilization of the ordinary least square method~\cite{leng2007ordinary} that is in charge for an efficient finding suitable coefficients of linearised formulas.  \fxnote{TODO: maybe expand the core of the algorithm... }

The main contributions of this work may be summarized as follows:

@@

	\item The new set of unbiased instances is introduced -- randomly generated formula of various sizes. All the other methods were tested against this new test-bed. The proposed \textsc{Rils-Rols} method showed to be dominant here as well. 
\end{enumerate}
